---
title: "Catch-Effort Figures"
author: "Camila Vargas"
date: "10/5/2020"
output: html_document
---

This script contains the functions to create catch-effort plot used in our project.
Note that we are plotting log(catch/effort) relationship and a moment of effort

There are three main plot/functions where you can choose the effort metirc to plot and the gear type besed on the GFW data.

- effort_moment_plot() - creates a one panel plot with all FAO areas
- effort_moment_fao_plot() - created a multiplanel plot, each pannel correspond to data for each FAO are
- qr_effort_catch_plot() - one panel plot with all FAO ares. Includes lines for 0.2, 0.5 and 0.8 quantile regressions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("R/set_up.R"))

##install and load packeges
load_pak(common_packages)
load_pak(plotting_pkgs)

##colors
area_cols <- 20
my_colors <- colorRampPalette(brewer.pal(8, "Set2"))(area_cols)
my_colors_2 <- colorRampPalette(brewer.pal(12, "Paired"))(area_cols)

```

##Read data

```{r}

catch_effort_data <- read_csv(here::here("output/catch_effort_long.csv"))

```


### Effort Moment plots

```{r}

effort_moment_plot <- function(gear_type, 
                               effort_metric, 
                               ordinate='slope_log', 
                               input_data = catch_effort_data){
  
  mplot <- input_data %>%
    filter(GFWCategory==gear_type, effort_type==effort_metric) %>%
    mutate(fao_area_code = as.factor(fao_area_code)) %>% 
    ggplot(aes(x = effort_moment, y = get(ordinate), label = iso3_code)) + ##label showa in the display windos when using ggplotly
      geom_point(aes(color = fao_area_code, size = n_vessel)) + #, shape = fao_area_code
      geom_text(aes(label = n_vessel), size = 2, check_overlap = T)+
      #scale_shape_manual(values=1:20)+
      #scale_color_manual(values = my_colors_2, name = "FAO Area Code")+
      geom_smooth(aes(weight=n_vessel), method = "lm", se = FALSE, linetype = "dotted")+
      geom_smooth(method = "lm", se = FALSE, linetype = "dotted", size=0.5)+
      #facet_wrap(~effort_type)+ #wrap vertically: ncol = 1
      theme_bw()+
      labs(title = paste0("Effort - Catch relationship for ", gear_type))+
      ylab(ordinate)+
      xlab(effort_metric)+
      theme(legend.position = "right")

  #ggplotly(mplot)
  return(mplot)
}

effort_moment_plot(gear_type = "seiners", effort_metric = "total_f_hours_tonnage")

effort_moment_plot(gear_type = "drifting_longlines", effort_metric = "total_f_hours_tonnage")

```


##Facet Plots

```{r}

effort_moment_fao_plot <- function(gear_type, 
                                   effort_metric, 
                                   input_data=catch_effort_data, 
                                   ordinate = "slope_log") {
 
  filt_data <- input_data %>%
    filter(GFWCategory == gear_type, 
           effort_type == effort_metric) %>%
    mutate(fao_area_code = as.factor(fao_area_code),
           ord = get(ordinate))
  
  model <- lm(ord ~ effort_moment, data=filt_data)
  
  fao_mplot <- filt_data %>% ggplot()+ 
      geom_smooth(method = "lm", se = TRUE, linetype = "dotted", # inherit.aes = FALSE,
                  data = model, aes(x = effort_moment, y = ord))+
      geom_point(aes(x = effort_moment, 
               y = ord, color = fao_area_code, size = n_vessel)) + #, shape = fao_area_code
      geom_text(aes(x = effort_moment, y = ord, label = n_vessel), size = 2, check_overlap = T)+
      scale_color_manual(values = my_colors_2, name = "FAO Area Code")+
      facet_wrap(~fao_area_code)+ #wrap vertically: ncol = 1
      theme_bw()+
      labs(title = paste0("Effort - Catch relationship for ", gear_type))+
      ylab(ordinate)+
      xlab(paste("moment of", effort_metric, sep=" "))+
      theme(legend.position = "right")

  #plot(fao_mplot)
  return(fao_mplot)

}

##By GEar type
#effort_moment_fao_plot(gear_type = "drifting_longlines", effort_metric = "total_f_hours_tonnage")

#effort_moment_fao_plot(gear_type = "seiners", effort_metric = "total_f_hours_tonnage")

#effort_moment_fao_plot(gear_type = "trawlers", effort_metric = "total_f_hours_tonnage")

#effort_moment_fao_plot(gear_type = "set_gillnets", effort_metric = "total_f_hours_tonnage")

```


## Quantile regression plots


```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.fullwidth=TRUE, fig.width=8, fig.height=6}


 qr_effort_catch_plot <- function(gear,
                                  data = catch_effort_data) {
   
   mplot <- data %>%
      filter(GFWCategory == gear) %>% 
      mutate(fao_area_code = as.factor(fao_area_code)) %>%
      ggplot(aes(x = effort_moment, 
                 y = slope_log, 
                 label = iso3_code)) + ##label show in the display window when using ggplotly
      geom_point(aes(color = fao_area_code, size = n_vessel)) + #, shape = fao_area_code
      geom_text(aes(label = n_vessel), size = 2, check_overlap = T)+
      scale_color_manual(values = my_colors_2, name = "FAO Area Code")+
      scale_size(name = "Number of Vessels")+
      #geom_smooth(aes(weight=n_vessel), method = "lm", se = FALSE, linetype = "dotted")+
      #geom_quantile(quantiles = c(0.8, 0.5, 0.2), aes(weight = n_vessel, colour = as.factor(..quantile..)))+
      geom_quantile(quantiles = 0.8, aes(color = "0.8"), color = "red")+
      geom_quantile(quantiles = 0.5, aes(color = "0.5"), color = "green")+
      geom_quantile(quantiles = 0.2, aes(color = "0.2"), color = "blue")+
      labs(color = "Quantile")+
      facet_wrap(~effort_type, scales = "free")+
      theme_bw()+
      #labs(title = gear)+
      labs(title = paste0( "<b><span style = 'font-size: 16pt'> Plot for ",
                           gear,
                       "</span></span></b>",
                       "</b><br> <span style = 'font-size:12pt'>",
                       "Quantile Regressions Lines ",
                       "<span style = 'color: red;'>",
                       "**0.8**, ",
                       "<span style = 'color:green;'>",
                       "**0.5**, ",
                       "</span></span></b>",
                       "and ",
                       "<span style = 'color:blue;'>",
                       "**0.2**",
                       "</span></span></b>"))+
    theme(plot.title = ggtext::element_textbox_simple(
      size = 13,
      face = NULL,
      lineheight = 1.75,
      padding = margin(5, 5, 0, 5),
      margin = margin(0, 0, 10, 0),#top,right,bottom,left.
      fill = "white")) +
   theme(plot.title.position = "plot",
         legend.position = "right")+
      ylab("log(effort/catch)")+ #
      xlab(paste("moment of", gsub('^.*_', '', "effort"), sep=" "))
  

   return(mplot)
   
    }

qr_effort_catch_regs <- function(gear,
                                  data = catch_effort_data,
                                 reg_color = "#33aadd") {
   
   mplot <- data %>%
      filter(GFWCategory == gear) %>% 
      mutate(fao_area_code = as.factor(fao_area_code)) %>%
      ggplot(aes(x = effort_moment, 
                 y = slope_log, 
                 label = iso3_code)) + ##label show in the display window when using ggplotly
      geom_point(aes(color = fao_area_code, size = n_vessel)) + #, shape = fao_area_code
      geom_text(aes(label = n_vessel), size = 2, check_overlap = T)+
      scale_color_manual(values = my_colors_2, name = "FAO Area Code")+
      scale_size(name = "Number of Vessels")+
      #geom_smooth(aes(weight=n_vessel), method = "lm", se = FALSE, linetype = "dotted")+
      geom_quantile(quantiles = seq(0.8, 0.2, by = -0.1), color = reg_color, show.legend = FALSE)+  
      geom_quantile(quantiles = 0.9, color = reg_color, linetype= "dashed", show.legend = FALSE)+  
           #labs(color = "Quantile")+
      #facet_wrap(~effort_type, scales = "free")+
     facet_wrap(~fao_area_code, scales = "free")+
      theme_bw()+
      #labs(title = paste("effort intensity by moment of effort,", gear, sep=" "))+
     labs(title = gear)+
    theme(plot.title = ggtext::element_textbox_simple(
      size = 13,
      face = NULL,
      lineheight = 1.75,
      padding = margin(5, 5, 0, 5),
      margin = margin(0, 0, 10, 0),#top,right,bottom,left.
      fill = "white")) +
   theme(plot.title.position = "plot",
         legend.position = "right")+
     guides(col = guide_legend(ncol = 2))+
      ylab("log(effort/catch)")+ #
      xlab("moment of effort")
  

   return(mplot)
   
    }


for (k in c("seiners", "trawlers", "drifting_longlines")) {
  ggsave(here::here(paste("output/figures/fig2-effort-intensity-", k, ".eps", sep="")), qr_effort_catch_regs(k), device="eps", width=320, height=160, units="mm", dpi="print", scale=0.8)
}


for (k in c("seiners", "trawlers", "drifting_longlines")) {
  ggsave(here::here(paste("output/figures/fig2-ei-compact-", k, ".eps", sep="")), qr_effort_catch_regs(k), device="eps", width=320, height=130, units="mm", dpi="print", scale=0.8)
}

tr <- qr_effort_catch_regs(gear = "trawlers")

ggplotly(tr)

qr_effort_catch_regs(gear = "seiners")

qr_effort_catch_regs(gear = "drifting_longlines")

```

