---
title: "R Notebook"
output: html_notebook
---

Setup

```{r}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("R/set_up.R"))

##install and load packeges
load_pak(common_packages)
load_pak(plotting_pkgs)

##colors
area_cols <- 20
my_colors <- colorRampPalette(brewer.pal(8, "Set2"))(area_cols)
my_colors_2 <- colorRampPalette(brewer.pal(12, "Paired"))(area_cols)

```

Load data

```{r}
catch_effort_data <- read_csv(here::here("output/catch_effort_long.csv"))

```

Plotting function

```{r}
qr_effort_catch_vers <- function(data,
                                 reg_color = "#33aadd") {
   
   mplot <- data %>%
      #filter(GFWCategory == gear) %>% 
      mutate(fao_area_code = as.factor(fao_area_code)) %>%
      ggplot(aes(x = effort_moment, 
                 y = slope_log, 
                 label = iso3_code)) + ##label show in the display window when using ggplotly
      geom_point(aes(color = fao_area_code, size = n_vessel)) + #, shape = fao_area_code
      geom_text(aes(label = n_vessel), size = 2, check_overlap = T)+
      scale_color_manual(values = my_colors_2, name = "FAO Area Code")+
      scale_size(name = "Number of Vessels")+
      #geom_smooth(aes(weight=n_vessel), method = "lm", se = FALSE, linetype = "dotted")+
      geom_quantile(quantiles = seq(0.8, 0.2, by = -0.1), color = reg_color, show.legend = FALSE)+  
      geom_quantile(quantiles = 0.9, color = reg_color, linetype= "dashed", show.legend = FALSE)+  
           #labs(color = "Quantile")+
      facet_wrap(~effort_type, scales = "free")+
      theme_bw()+
      #labs(title = paste("effort intensity by moment of effort,", gear, sep=" "))+
     #labs(title = GFWCategory)+
    theme(plot.title = ggtext::element_textbox_simple(
      size = 13,
      face = NULL,
      lineheight = 1.75,
      padding = margin(5, 5, 0, 5),
      margin = margin(0, 0, 10, 0),#top,right,bottom,left.
      fill = "white"), strip.text.x = element_text(size=10)) +
   theme(plot.title.position = "plot",
         legend.position = "right")+
     guides(col = guide_legend(ncol = 2))+
      ylab("log(effort/catch)")+ #
      xlab("moment of effort")
  

   return(mplot)
   
    }

```

For this plot, we want two exemplary catch-effort regressions for two different gear types.
We also want to show only data we actually used, which was as follows:
 - longliners, length
 - seiners, tonnage
 - trawlers, length
 

I say we do longliners length, seiners tonnage.  So we do the two filtrations, then
we stick them together, and we run with the facet-wrap on effort metric- apart from
some axis labeling it should take care of itself.

```{r}
lll <- catch_effort_data %>% filter(GFWCategory == 'drifting_longlines', 
                                    effort_type == 'total_f_hours_length') %>%
   mutate(effort_type = recode(effort_type, "total_f_hours_length" = "longlines, LOA"))
   
st <- catch_effort_data %>% filter(GFWCategory == 'seiners',
                                   effort_type == 'total_f_hours_tonnage') %>%
   mutate(effort_type = recode(effort_type, "total_f_hours_tonnage" = "seiners, GT"))

tl <- catch_effort_data %>% filter(GFWCategory == 'trawlers', 
                                   effort_type == 'total_f_hours_length',
                                   iso3_code != 'CHN') %>%
   mutate(effort_type = recode(effort_type, "total_f_hours_length" = "trawlers, LOA")) 

both <- rbind(lll, st)

reg_plot <- qr_effort_catch_vers(both)

allthree <- rbind(tl, st, lll)

r3_labels <- data.frame(effort_type = c('longlines, LOA', 'seiners, GT', 'trawlers, LOA'), 
                        label = c('A', 'B', 'C'))

reg3_plot <- qr_effort_catch_vers(allthree) + 
   geom_text(aes(x=-Inf, y=Inf, label = label), data=r3_labels, hjust = -1, vjust=2, size = 8)

```

Then I need to figure out how to automatically export to a file- but for now I did it manually, resizing to 1201x554 px and exporting as EPS to output/figures/effort_moment_regression_3panel_1201_554.eps

