---
title: "Catch Effort Quantile Regressions"
author: "Camila Vargas"
date: "11/5/2020"
output: pdf_document
---

This scrip contains the quantile regression anaysis for the log10(catch/effort) realationship with effort moment.
We calcilate coeficcients and p-values for trawlers, purse seiners and drifting longlines. However, the general functions `qr_calculations` and `qr_pvalue` can be applies to any gear in the GFW data.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("R/set_up.R"))

##install and load packeges
load_pak(common_packages)

load_pak(additional_pkgs)

```

## Data

```{r, include=FALSE, echo=FALSE}

##Raw data
catch_effort_data <- read_csv(here::here("output/catch_effort_long.csv"))

```

## Functions to run quantile regression and extract coefficients and p-values

```{r, echo=FALSE, include=FALSE}


tau_values <- c(0.2, 0.5, 0.6, 0.7, 0.8, 0.9)


select_gears <- function(gear_type, effort_metric, dataset = catch_effort_data){
  
  return(dataset %>% 
          filter(GFWCategory == gear_type,
         effort_type == effort_metric,
         !is.na(slope_log | effort_moment)) %>% 
  select(year, iso3_code, GFWCategory, fao_area_code, n_vessel,effort_type, slope_log, effort_moment))
}


qr_ffcalculations <- function(gear_type, effort_metric, dataset = catch_effort_data){
  # Do not weight by n_vessel because data for small vessels is unreliable
  
  catch_effort_gear <- select_gears(gear_type, effort_metric, dataset)
 
  gear_weight_reg <- rq(slope_log ~ effort_moment, tau = tau_values, data = catch_effort_gear)  #, weights = n_vessel) 
  
  reg_r <- tidy(gear_weight_reg, se = "nid") %>%  # note that specifying se = "nid" automatically appends standard error data
    mutate(gear = gear_type,
           effort_type = effort_metric)
  return(reg_r)
}


```

## Calculations for each combination

```{r, echo=FALSE, include=FALSE}

##Dfrinting Longlines
dll_length <- qr_ffcalculations(gear_type = "drifting_longlines",
                              effort_metric = "total_f_hours_length")

dll_tonnage <- qr_ffcalculations(gear_type = "drifting_longlines",
                              effort_metric = "total_f_hours_tonnage")

dll_kw <- qr_ffcalculations(gear_type = "drifting_longlines",
                              effort_metric = "total_f_hours_kw")


##Seiners
seiners_length <- qr_ffcalculations(gear_type = "seiners",
                              effort_metric = "total_f_hours_length")

seiners_tonnage <- qr_ffcalculations(gear_type = "seiners",
                              effort_metric = "total_f_hours_tonnage")

seiners_kw <- qr_ffcalculations(gear_type = "seiners",
                              effort_metric = "total_f_hours_kw")

##trawlers

# For Trawlers, when weighting by n_vessel (no longer recommended), it is 
# necessary to filter out China prior to fitting the regression because their 
# n_vessel ~= 16,000 which is far larger than any other country.

# catch_effort_no_chn <- catch_effort_data %>%  filter(iso3_code != 'CHN')

trawl_length <- qr_ffcalculations(gear_type = "trawlers",
                              effort_metric = "total_f_hours_length", 
                              dataset = catch_effort_data)

trawl_tonnage <- qr_ffcalculations(gear_type = "trawlers",
                              effort_metric = "total_f_hours_tonnage", 
                              dataset = catch_effort_data)

trawl_kw <- qr_ffcalculations(gear_type = "trawlers",
                              effort_metric = "total_f_hours_kw", 
                              dataset = catch_effort_data)

## ALL TOGETHER

all_reg_results <- rbind(dll_length, dll_tonnage, dll_kw, seiners_length, seiners_tonnage, seiners_kw, trawl_length, trawl_tonnage, trawl_kw) %>%
  clean_names() %>% 
  mutate(term = case_when(term == "effort_moment" ~ "coefficient",
                          T ~ "intercept"))

#all_reg_pvalue <- rbind(dll_length_p, dll_tonnage_p, dll_kw_p, seiners_length_p, seiners_tonnage_p, seiners_kw_p, trawl_length_p, #trawl_tonnage_p, trawl_kw_p) %>% 
#  clean_names()


#reg_results <- all_reg_results %>% 
#  left_join(all_reg_pvalue, by = c("value", "tau", "gear", "effort_type"))



```

## Save results

```{r}
write_csv(all_reg_results, here::here('output/quantile_reg_results.csv'))

```



