---
title: "Master Tables"
author: "Camila Vargas"
date: "5/20/2020"
output: html_document
---


This script creates the countries' master table to match countries and from the Watson's data and FAO fisheries landings and the corresponding the ISO3 code.
Throughout the analysis, the ISO3 code is used to identify a country. 

Note: countries with no ISO3 code
|---|---|---|
| 1| 058  |   Belgium-Luxembourg      |
| 2| 128  |   Canton and Enderbury Is |
| 3| 230  |   Ethiopia PDR            |
| 4| 274  |   Gaza Strip(Palestine) |  
| 5| 278  |   Germany,New Länder      |
| 6| 280  |   Germany, Fed. Rep. of   |
| 7| 402  |   West Bank               |
| 8| 536  |   Neutral Zone            |
| 9| 582  |   Pacific Islands Trust Tr|
|10| 680 |    Sark                    |
|11| 720 |    Yemen, Democratic       |
|12| 736 |    Sudan (former)          |
|13| 830 |    Channel Islands         |
|14| 886 |    Yemen Arab Republic|     
|15| 896 |    Other nei   |

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("R/set_up.R"))

##install and load packeges
load_pak(common_packages)

```


##Watson Country Codes

The Watson data base uses number ID to identify countries. They provide a key to match countries ID with names, however those names do not match the names on the FAO landing data base.

```{r}

watson_codes_country <- read_excel(here::here("data/watson/watson_codes.xlsx"), sheet = "Country") %>% 
  clean_names() %>% 
  mutate(fao_name = str_trim(fao_name, side = "both")) %>% 
  rename(watson_country_id = country)

###
##write_csv(watson_codes_country, here::here("mapping_tables/int/watson_codes_country.csv"))
```

We exported the Watson country codes and manually match these names with the FAO country name and the FAO landings country name.


##Importing the csv with the matching names

Note: solving encoding isssue by creating the fao_test column. However that makes all the names encoded as UTF-8 to become NAs. To solve this second issue we created a 3red column, fao_landing_name.

```{r}


watson_to_fao <- read_xlsx(here::here("mapping_tables/int/watson_fao_country_key.xlsx")) %>%
  mutate(fao_test = ifelse(watson_c_name == "Br Virgin Is","BVI", fao_landing_c_name)) %>% 
  mutate(fao_landing_name = ifelse(is.na(fao_test), fao_landing_c_name, fao_test)) %>% 
  select(-fao_landing_c_name, -fao_test) %>% 
  rename(fao_landing_c_name = fao_landing_name) %>% 
  mutate(fao_landing_c_name = case_when(
    fao_landing_c_name == "BVI" ~ "British Virgin Islands",
    fao_landing_c_name == "French Polynisia" ~ "French Polynesia",
    T ~ fao_landing_c_name))

```


##Combining watson to FAO with fao_countries

Finally, to fill in the missing the ISO names and the FAO official names, we use the CL_FI_COUNTRY_GROUPS.csv file. This files comes from the capture production zip folder downloaded form the (FAO web site)[http://www.fao.org/fishery/statistics/global-capture-production/en] 
Downloaded on May 26, 2020 and saved in the `fao_landings` folder under the name `fao_country_names_groups.csv`

We combine these two data set to have all the countries names and abreviations in one master table.

The final master table is saved as a csv file!

```{r}

fao_country_key_complete <- read_csv(here::here("data/fao_landings/fao_country_names_groups.csv"))%>% 
  clean_names() %>% 
  select(1:5, 11) %>% ##Selecting only english names and codes of our interest
  rename(fao_landing_c_name = name_en)
  
master_countries <- fao_country_key_complete %>%
  left_join(watson_to_fao, by = "fao_landing_c_name")
  
#write_csv(master_countries, here::here("mapping_tables/master_countries.csv"))
```


**NOTE:**
- The FAO landings data base has data for more countries than the Watson data base. Probably some of the differences are on countries that report inland waters landings (landlock countries)



##Table with FAO names from FAO website

```{r}
##Additional packages to retrived data from the web
# install.packages("rvest")
# library(rvest)
```


This is a list of all the countries that participate in the FAO. Note that not all of them have fishing activity.

Note: Theses names not necesarly match the Counties name in the landings data
Note 2: This table was an initial approach so retrive the FAO country names. We ended up not using this infomation but the file below.

```{r}
# fao_website_iso <- read_html("http://www.fao.org/countryprofiles/iso3list/en/")
# 
# tbls <- html_nodes(fao_website_iso, "table") %>% 
#   html_table(fill = TRUE)
# 
# fao_countries <- as.data.frame(do.call(cbind, tbls)) %>% 
#   clean_names() %>% 
#   select(1:4) %>% 
#   mutate(fao_name = str_remove(short_name, "(?<=\\().+")) %>% 
#   mutate(fao_name = str_remove(fao_name, "\\(")) %>% 
#   mutate(fao_name = str_trim(fao_name, side = "both"))
  
```
